An improved metric for the analysis of swarms

### 2017-04-05 paper.tex
Magnitude Updates

\section{Magnitude based metric\label{section:MagnitudeDynamics}}
Magnitude based internal movement (\textit{agent resultant magnitude}) is measured by identifying the balance between the repulsion and cohesion between agents. `Jitter' in the case of the \textit{agent resultant magnitude} metric is measured as the variance of the magnitude lengths created by the agents. The identification of this variance produces the clarifying part of the \textit{agent resultant magnitude} metric.
%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
The \textit{agent resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. There are two ways of using the cohesion and repulsion in identifying a resultant vector. The two vectors can be added as absolute values to give an overall `size' to the magnitude that is affecting each relationship. Alternatively the resultant magnitude can be the sum of the actual magnitudes. The repulsion vector has a negative magnitude and the cohesion vector has a positive magnitude. In this paper the magnitude analysis will be based on summing the two actual vectors to determine the result of the inter-agent interaction, although both mechanisms could be used for a comparitive analysis. This paper will refer to the resultant magnitude as the \text{`agent resultant magnitude'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. the \text{`agent resultant magnitude'} can be considered part of the `quality' measure for a swarm's performance.

If the \textit{agent resultant magnitude} is a negative value (absolute values would prevent this analysis) the swarm's bias is to expand. This is seen in the chaotic stage of a swarm. If the \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity. This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.

The \textit{agent resultant magnitude} on its own does not give a complete measure of a swarm's internal state. There needs to be a qualifying component to the metric  that identifies the degree of deviation in the resultant magnitude, this is the \textit{jitter}. The smaller the degree of deviation the more uniform the structure of the swarm. These two components identify the degree to which a swarm has progressed towards a stable state.

The \textit{agent resultant magnitude} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent. The variance component identifies the degree to which the swarm has stabilised. The ideal status for inter-agent interactions would be for the agents to have a resultant vector (\textit{agent resultant magnitude}) of zero or above. This would indicate that the agents are distributed such that they are at their distribution limit (outer most range of the cohesion field) or at a level that causes the agents to `pull' together. The ideal degree of deviation is zero as this indicates an even distribution of agents. Therefore for a fuller indication of a swarms \textit{state} both measurements need to be combined. The deviation from the mean clarifying the internal movement and the \textit{agent resultant magnitude} providing an indication of the `compression' that a swarm is logically experiencing (cohesiveness). These two aspects of a swarm's features are not considered by Gazi and Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying the structure of a swarm in terms of stability.

\section{Magnitude based internal movement model}\label{Section:StabilityModel}
Using the formulae for the calculation of cohesion~(Equation \ref{eq:FlyToCentre1}, page \pageref{eq:FlyToCentre1}) and repulsion~(Equation \ref{eq:Repulsion1}, page \pageref{eq:Repulsion1}) for every agent and its neighbours it is possible to calculate an \textit{agent resultant magnitude} value (sum of agent resultant magnitudes). This value represents the overall potential of an agent. This magnitude when normalised produces a component of the \textit{movement-destination vector} for a swarm. If the agent resultant magnitude is zero (null vector) then the agent will not move. $P(b)$ is the \textit{inter-agent resultant magnitude vector} for agent $b$ defined by:

@@

To identify the swarm based \textit{agent resultant magnitude}~\ref{eq:SwarmStabilityMetricT} must be extended to iterate over all the agents in the swarm. \ref{eq:SwarmStabilityMetricT3} shows $\mu_p(S)$~as the swarm based magnitude where the swarm iteration is shown as~$\sum_{b \in S}$ and $\sum_{b \in S}|nbr(b)|$ calculates the total number of inter-agent relationships.

@@

\section{Variance in agent resultant magnitude metric}\label{Section:VarianceInPotential}
The mechanism just described provides an overall indication of the internal movement based on inter agent vectors that produce the \textit{agent resultant magnitude}. This model however is not sufficient to give an indication of the swarm `state' as an overall metric. To improve the metric clarification is required in terms of the deviation from the \textit{agent resultant magnitude} norm. The variation in the metric is the standard deviation of the entire swarm from the mean of the inter-agent potential magnitudes (Equation \ref{eq:SwarmStabilityMetricT}).

The standard deviation is calculated as Equation~\ref{eq:SwarmStabilityQuotientT} where $\sigma_p(S)$ is the standard deviation at a time $t$ and $\mu_p(S)$ is the mean at the same point in time. $\sum_{b \in S}~\sum_{b' \in nbr(b)}$ iterates over every agent in the swarm and its neighbours and $\sum_{b \in S}|nbr(b)|$ calculates the total number of inter-agent relationships.

\begin{equation}
\label{eq:SwarmStabilityQuotientT}
\sigma_p(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b')}}\Big(P(b')-\mu_p(S)\Big)^2}{\mathlarger{\sum_{b \in S}}|nbr(b))|}}
\end{equation}

The metric for the internal movement is a set of numbers, the mean and standard deviation of the swarms internal \textit{agent resultant magnitude} derived from each agent and its neighbour interactions~(\ref{eq:SwarmPotentialMagnitude}). The pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:

@@

Both metrics identify the state of a swarm with respect to variations in the dispersement of the agents from an average distribution.

@@

The \textit{agent resultant magnitude} metric provides a view of the interaction magnitude. This provides an indication of the swarms potential movement. This is independent of the physical distribution. The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~\ref{additional:fieldsWork} where the physical distribution may vary.

>>>>

\section{Magnitude based metric\label{section:MagnitudeDynamics}}
Magnitude based internal movement (\textit{agent resultant magnitude}) is measured by identifying the balance between the repulsion and cohesion between agents. `Jitter' in the case of the \textit{agent resultant magnitude} metric is measured as the deviation of the magnitude lengths created by the agents. The identification of this deviation produces the clarifying part of the \textit{agent resultant magnitude} metric.
%% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
The \textit{agent resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. There are two ways of using the cohesion and repulsion in identifying a resultant vector. The two vectors can be added as absolute values to give an overall `size' to the magnitude that is affecting each relationship. Alternatively the resultant magnitude can be the sum of the actual magnitudes. The repulsion vector has a negative magnitude and the cohesion vector has a positive magnitude. In this paper the magnitude analysis will be based on summing the two vectors to determine the result of the inter-agent interaction, although both mechanisms could be used for a comparitive analysis. This paper will refer to the resultant magnitude as the \text{`agent resultant magnitude'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. the \text{`agent resultant magnitude'} can be considered part of the `quality' measure for a swarm's performance.

If the \textit{agent resultant magnitude} is negative (absolute values would prevent this analysis) the swarm's bias is to expand as the repulsion is greater than cohesion. This is seen in the chaotic stage of a swarm. If the \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity. This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.

The \textit{agent resultant magnitude} on its own does not give a complete measure of a swarm's internal state. There needs to be a qualifying component to the metric  that identifies the degree of deviation in the resultant magnitude, this is the \textit{jitter}. The smaller the deviation the more uniform the structure of the swarm. These two components identify the degree to which a swarm has progressed towards a stable state.

The \textit{agent resultant magnitude} provides a view of the swarm's state through the balance between the repulsive and the cohesive vectors that are being applied to each agent. The deviation component identifies the degree to which the swarm has stabilised. The ideal status for inter-agent interactions would be for the agents to have a resultant vector (\textit{agent resultant magnitude}) of zero or above. This would indicate that the agents are distributed such that they are at their distribution limit (outer most range of the cohesion field) or at a level that causes the agents to `pull' together. The ideal degree of deviation is zero as this indicates an even distribution of agents. These two aspects of a swarm's features are not considered by Gazi and Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying the structure of a swarm in terms of stability.

\section{Magnitude based internal movement model}\label{Section:StabilityModel}
Using the formulae for the calculation of cohesion~(Equation \ref{eq:FlyToCentre1}) and repulsion~(Equation \ref{eq:Repulsion1}) for every agent and its neighbours it is possible to calculate an \textit{agent resultant magnitude} value (sum of agent resultant magnitudes). This value represents the overall `potential' of an agent. This magnitude when normalised produces a component of the \textit{movement-destination vector} for a swarm. If the agent resultant magnitude is zero (null vector) then the agent will not move. $P(b)$ is the \textit{inter-agent resultant magnitude vector} for agent $b$ defined by:

@@

To identify the swarm based \textit{agent resultant magnitude}~Equation~\ref{eq:SwarmStabilityMetricT} must be extended to iterate over all the agents in the swarm. \ref{eq:SwarmStabilityMetricT3} shows $\mu_p(S)$~as the swarm based magnitude.

@@

\section{Standard deviation in agent resultant magnitude metric}\label{Section:VarianceInPotential}
To improve the metric clarification in terms of the deviation from the \textit{agent resultant magnitude} norm is calculated. The standard deviation of the entire swarm  is shown in Equation \ref{eq:SwarmStabilityMetricT}.

The standard deviation is calculated as Equation~\ref{eq:SwarmStabilityQuotientT} where $\sigma_p(S)$ is the standard deviation at a time $t$ and $\mu_p(S)$ is the mean at the same point in time.

\begin{equation}
\label{eq:SwarmStabilityQuotientT}
\sigma_p(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b)}}\Big(P(b')-\mu_p(S)\Big)^2}{\mathlarger{\sum_{b \in S}}|nbr(b))|}}
\end{equation}

The metric for the internal movement is a set of numbers, the mean and standard deviation of the swarm's internal \textit{agent resultant magnitude} derived from each agent and its neighbour interactions~(Equation~\ref{eq:SwarmPotentialMagnitude}). The pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:

@@

Both metrics identify the state of a swarm with respect to deviations in the dispersement of the agents from a mean.

@@

The \textit{agent resultant magnitude} metric provides a view of the interaction magnitude. This provides an indication of the swarm's potential movement. This is independent of the physical distribution. The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~\ref{additional:fieldsWork} where the physical distribution of agents may vary but due to the differing feild ranges the resultant magnitudes balance.
